Autonomous Oceanographic Data Collection towards Estimating Excursion Sets of Vector-valued Random Fields

### 2020-06-12 ans_to_reviewers/answers-to-reviewers.tex

% == Maths
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[normalem]{ulem} %to strike the words

\newcounter{reviewer}
\def\revcom{\textbf{Reviewer comment}}
\def\aecom{\textbf{Associate Editor comment }}
\def\reply{\textbf{Reply}}
\def\action{\textbf{Action}}

\newlist{answers}{enumerate}{10}
\setlist[answers]{leftmargin=*, label=\arabic{reviewer}.\arabic*}

\newcommand{\edcomment}[1]{{\color{green}{\{Editor: 1\}}}}
\newcommand{\frevcomment}[1]{{\color{blue}{\{Rev 1: 1\}}}}
\newcommand{\srevcomment}[1]{{\color{red}{\{Rev 2: 1\}}}}
\newcommand{\trevcomment}[1]{{\color{violet}{\{Rev 3: 1\}}}}

@@

We are grateful to the reviewers for their comments and
suggestions. They have all been considered in this revised
version of our manuscript. We have worked a lot on developing both theoretical results, computational resources, and on the presentation in the paper. To help us in these efforts, we included Cedric Traveletti as co-author because he has contributed significantly on the revised version of the paper.

The principal revisions made to the manuscript are as shown in this list, and they are done to address the main comments of Editors and Reviewers:

\begin{itemize}
\item The motivation for excursion sets (ESs), and in particular that of {\bf{multivariate ESs}}, has been strengthened. Motivating examples are abundant in medicine where doctors do not solely rely on a single symptom but must see several combined effects before making a diagnosis. Our motivation in the current paper is that temperature and salinity are proxies for more complex biological variables in the ocean, and in several cases they must be jointly high or low for phenomena to occur. For instance, salmon lice larvae are known to depend on relatively high temperature and high salinity to survive. Salmon lice are parasites to salmon, and the bivariate ES for temperature and salinity is hence important to understand for sustainable activities in salmon fisheries.

In the revised manuscript these aspects are largely rewritten in the Introduction, and partly also in Section 2 on the expected integrated Bernoulli variance (EIBV) and partly in the Case Study ({\bf{TO DO 6 AGAIN}}). This change addresses some key comments of AE, Rev 1 in particular.

@@

\item The {\bf{statistical methodological contribution}} have been clarified and improved, including definitions of random sets and ESs, and having more general results than just the EIBV for the bi-variate case of temperature and salinity. The revised version highlights these novel statistical results as a main contribution of the paper.

Moreover, we emphasize the statistical design problem which is sought solved by AUV with its constraints on navigation and computing resources.

In this sense, the revised version should provide ample motivation for statisticians, while still maintaining strong links to the application for illustrating concepts.

In the revised manuscript these aspects are largely rewritten in the Introduction, in Sect 2 and 3, and on future work in the Closing remarks. The main body contains the novel statistical results for the expected uncertainty reduction in random sets exemplified by the EIBV, while the Appendix holds the proofs of key results. This change addresses the some comments of Editor, Associate Editor, Rev 1, Rev 2 and Rev 3.

\item We realized that a contribution of the current work is available code for practitioners or statisticians who could try this themselves, or possibly extend it for their own cases. We have hence made a substantial effort in structuring {\bf{computer code for examples}} which are presented on a URL link. (\url{https://cedrictravelletti.github.io/Submarine/} )
 to go with the paper. This open-source python code is also ready to run on more nuanced statistical models (non-stationary and non-isotropic).

The simulation study section contains pointers to the associated open-source code. This change addresses the comments of Rev 1 and Rev 2 in particular.

@@

%\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]

%\item[1.1.1]\label{ec1} 
\vspace{5mm}
\noindent \aecom \textbf{1}:

General: The inferential problem considered is the paper is that of estimating the “boundary” of a plume. However, the authors never define exactly what a “boundary” is. Such a definition should be provided, as well as some discussion of why determining the location of a boundary is so important. Why is it not sufficient to create a map of predictions (with an accompanying map of prediction error variances) for the variable(s) under study throughout the region of interest (a problem for which there is a huge literature on sampling design)?\par

\reply: The revised version has an entirely new Section 2 which includes general background for random sets and excursion sets. We added references, for instance \cite{adler2009random} on excursion sets and \cite{Molchanov2005} on random sets.

@@

\vspace{1em}

%\item[1.1.3]\label{ec3} 
\vspace{5mm}
\noindent \aecom \textbf{3}:

Page 9, second paragraph of Section 4.2: It has been assumed here that the measurement errors for temperature and salinity at any given location are uncorrelated. Is this a reasonable assumption? This could easily be checked by taking replicate measurements at each site.\par

\reply: The assumption is that salinity and temperature are conditionally independent measurements. (In the latent Gaussian random field, the variables are dependent.) We believe this conditional independence assumption is realistic as the measurements are gathered from different sensors. We added a comment about this not necessarily being the case with other data that are extracted and processed by combining sensors. Then assumptions can be tested in a lab under idealized conditions. But it is difficult to do so for the AUV in the changing oceans.

@@

%\item[1.2.1]\label{emc1} 
\item Page 3, line 19: Change “bi-variate” to “bivariate”. The same change is needed on the next-
to-last line of Page 12.

\reply: We have changed this to be consistent in the revised text.

%\item[1.2.2]\label{emc2} 
\item Page 9, three lines after (10): I presume that all off-diagonal elements of R are zeros, but this
should be stated explicitly.

\reply: We have changed the statement on page 9 to: ``$\bR$ contains zeroes except on the diagonal entries, which are set to $r^2_1$ and $r^2_2$, holding the measurement error variances of temperature and salinity observations."

@@

%\item[1.2.3]\label{emc3} 
\item Page 19, fourth line from bottom: Change ``as also" to ``as well as"

\reply: We have changed this in the text - page 19.

%\item[1.2.4]\label{emc4} 
\item Page 30, second and third references listed: ``et al." is listed after the authors’ names for each
of these two references, but actually there are no additional authors beyond the two listed.

\reply: We have made this change to the citations: \cite{genton2015cross} and \cite{french2016credible} - page 30.
\end{itemize}

@@

\reply: We incorporated the suggested minor changes.

\item{\revcom  :}\label{r1c4}

I am not clear about the intended audience for this paper. Or to put it another way, for whom is this paper important? The case study or the
design methodology? Case studies offer concern specific events deemed of importance e.g. the much–studied O-ring failure long ago on a US rocket launch. Such a case study may then drive the innovation of statistical methods for an analysis of data that were collected because of that case’s importance. More likely the case study was intended as a tutorial on the use of a novel design methodology. In fact the Closing Remarks section points “the opportunities available real-time multivariable spatial data-gathering” that “statistician can exploit”.

\reply: The motivation for an applied statistical audience is strengthened, while we have maintained strong links to the applications of Robotics, Ocean Science and AI. In particular in the Introduction and in Section 2 are now stronger on motivating multivariate ESs, and highlighting the need for new statistics in these marine robotics applications. See main comments early on in this reply letter.

@@

However, the user community is pretty tech-savvy nowadays, and that community might well be the primary beneficiary and more motivated than the statisticians to use the tools for sequentially sampling rivers for mapping plumes. That suggests including the relevant code for running the analyses, especially the simulation studies. An alterna- tive journal to a statistical one might be more appropriate from that perspective.

\reply: We have made an open-source repository available for the public. This includes commented code examples for exploration and analysis of the material presented in the paper. The code is optimized for speed to allow fast computation using Pytorch (citation). The code is available here: (\url{https://cedrictravelletti.github.io/Submarine/} )

@@

\reply: TO DO

\item{\revcom :}\label{r2c14}

While the context does not allow full reproducibility of the paper results, no implementation details are provided. What software was used to perform the bulk of analysis? Are there scripts available that can be adapted by others for similar purposes? Please provide resources for the practical implementation of the methodology discussed in the paper.

\reply: Code is improved in the revision and provided for open-source usage (\url{https://cedrictravelletti.github.io/Submarine/} )

@@

\reply: This is changed in the revised version.

\item{\revcom - Major 2:}\label{r3c3}

Figure 4: Why not show excursion probabilities for, say, myopic selection as well?

\reply: In the revised version we show EPs for a static survey line, myopic and look-ahead selection in three displays. (This is in the new Figure 3.)

In the accompanying code (\url{https://cedrictravelletti.github.io/Submarine/} )
there are also resources for displaying other outputs if the reader is interested in comparing results with for instance univariate excursion probabilities.

### 2020-06-12 UQonSets.tex
Updates from Overleaf

Let us in particular consider a locally finite (Borel) measure $\mes$ on $D$ and investigate the probability distribution 
of $\mes(\es)$ through its moments.

@@

%That $\mu(\es)$ is a well-defined random variable follows from
\begin{propo}
    \label{propo1}
%Under technical conditions to be precised, $\mu(\es)$ possesses moments or arbitrary order and these can be expressed for any $r\geq 1$ as
For a measurable random field $\gp$ and a locally finite measure $\mes$ on $D$, $\mes(\es)$ is a random variable and for 
any $r\geq 1$,
\begin{equation*}
\begin{split}
\mathbb{E}[\mes(\es)^r]
&=\int_{D^{r}} \jointExcuProb
\productMeasure
,
\end{split}
\end{equation*}

where the product measure is denoted as
$\nu^{\otimes}:=\bigotimes_{i=1}^r \nu$.
%Note that we will use boldface to denote concatenated variables. % TO RELOCATE
Here $\gp$ is defined on $D$, and for
$\bm{u}=\left(u^{(1)}, ..., u^{(r)}\right)\in D^r$, $\gp[\bm{u}]=\left(\gp[u^{(1)}], ...,
\gp[u^{(r)}]\right)\in \mathbb{R}^{\no r}$.
\medskip

In the particular case where $\gp$ is a multivariate Gaussian random field
%and denoting the mean and covariance matrix of $\gp[\bm{u}]$ by $\meanUU\in\mathbb{R}^{p\times r}$ and $\covUU$, respectively,
we have
\begin{align*}
\jointExcuProb = \mathcal{N}_{\no r}(T^r; \meanUU, \covUU),
\end{align*}
where $\mathcal{N}_{\no r}(\cdot ; \meanUU, \covUU)$ is the Gaussian measure on $\mathbb{R}^{\no r}$ with mean $\meanUU$ 
and covariance matrix $\covUU$, respectively defined blockwise by
%where both are defined blockwise by
\begin{align*}
\meanUU&=\begin{pmatrix}\mu(u^{(1)})\\ \vdots\\ \mu(u^{(r)})\end{pmatrix}
\in \mathbb{R}^{\no r}, \\
\text{and } \covUU &= \begin{pmatrix}
\cov(\gp[u^{(1)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(1)}],
\gp[u^{(r)}])\\
\vdots & & \vdots\\
\cov(\gp[u^{(r)}], \gp[u^{(1)}]) & \dots & \cov(\gp[u^{(r)}],
\gp[u^{(r)}])\\
\end{pmatrix}\in \mathbb{R}^{pr\times pr},
\end{align*}
each of the $r\times r$ blocks of the latter matrix being itself a (cross-)covariance matrix of dimension $\no \times 
\no$.
%dimensional Gaussian random vector.
Assuming further that $\covUU$ is non-singular, the probability of interest can be formulated in terms of the $\no 
r$-dimensional Gaussian probability density function
$\varphi_{\no  r}(\cdot;~\meanUU, \covUU)$ as
%$\varphi_{\no \times r}$ as
%$\bm{u}\in D^r$ such that $\gp[\bm{u}]$ be non-degenerate, the integrand can be expressed in terms of the $r \times \no$-dimensional Gaussian distribution, namely
%cumulative distribution function, namely
\begin{equation*}
\begin{split}
&\jointExcuProb
=
\int_{T^r} \varphi_{\no  r}\left(\bm{v};
    ~\meanUU, \covUU\right)
    \mathrm{d}\bm{v},
\end{split}
\end{equation*}
In the particular orthant case with $\T=(-\infty, t_1] \times \dots \times (-\infty, t_{r}]$,
the latter probability directly writes in terms of the multivariate Gaussian
cumulative distribution, % with mean $\meanUU$ and covariance matrix $\covUU$,
this time by the way without requiring $\covUU$ %the latter
to be non-singular:
\begin{equation*}
\begin{split}
\jointExcuProb
&=
\varPhi_{\no r}\left(\bm{t};~\meanUU, \covUU\right),
\end{split}
\end{equation*}
where we have used the notations
$t=(t_1,\dots,t_{\no}%,..., t_1, ..., t_p
)\in\mathbb{R}^{\no}$, $1_{r}=(1,\dots,1)\in \R^{r}$, and
$\bm{t}=1_{r}\otimes \bm{t}=(t_1,\dots,t_{\no},\dots,t_1,\dots,t_{\no})
\in \R^{\no r}$.
\end{propo}
\begin{proof}
That $\mes(\es)$ defines indeed a random variable follows from Fubini's theorem
relying on the joint measurability of
$(\x, \omega) \to \mathbbm{1}_{\es(\omega)}(\x)$,
%= \mathbf{1}_{\xi_{\x}(\omega) \in T}$ and Fubini's theorem.
itself inherited from the assumed measurability for
$(\x, \omega) \to \gp[\x](\omega)$ and $T$, respectively. From there, following the steps of Robbins' theorem \cite{Robins1944}, we find that
\begin{equation*}
\begin{split}
\mathbb{E}[\mes(\es)^{r}]
&=\mathbb{E}\left[\left(\int_{D} \mathbbm{1}_{\gp[u] \in T} ~d\mes(u) \right)^{r} \right]
=\mathbb{E}\left[ \prod_{i=1}^{r} \left(
        \int_{D} \mathbbm{1}_{\gp[\uu^{(i)}] \in T} ~d\mes(\uu^{(i)})
\right) \right] \\
&
=
%\mathbb{E}\left[
%\int_{D^{r}}
%\prod_{i=1}^{r}
%\mathbf{1}_{\gp[\uu^{(i)}] \in T}
%\productMeasure
%\right]
%=
\mathbb{E}\left[
\int_{D^{r}}
\mathbbm{1}_{\gp[\uu^{(1)}] \in T,\dots, \gp[\uu^{(r)}]  \in T}
\productMeasure
\right]
%\\
%&=\int_{D^{r}} \mathbb{E}\left[\mathbf{1}_{\gp[\bm{u}] \in T^r} \right] \productMeasure \\
%&
=\int_{D^{r}}
\jointExcuProb
\productMeasure.
\end{split}
\end{equation*}
The rest consists in expliciting the probability of $T\times \dots \times T$ under the multivariate Gaussian distribution of
$\left(\gp[u^{(1)}], \dots,  \gp[u^{(r)}] \right)$.
\end{proof}

Proposition~\ref{propo1} enables in turn calculating centred moments of $\mes(\es)$.
The excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ is probably the most relevant of them for the considered purposes, and hence writes
\begin{equation*}
\begin{split}
\emv
&=\int_{D^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right)
d\mes^{\otimes}(u, v)\\
&-\left( \int_{D} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
\dots \times (-\infty, t_{\no}]$ to
\begin{equation*}
\begin{split}
\emv
%\operatorname{Var}[\mes(\es)]
&=\int_{D^2}
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)),
K((u,v),(u,v))
\right)
\
\mathrm{d}\mes^{\otimes} %\mes
%\productMeasure
(u,v)\\
&-\left( \int_{D} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
%
Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $D^2$. In 
contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves 
solely and integral on $D$ and can be expanded in our present settings as follows
\begin{equation*}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{D}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{D}
\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u).
\end{split}
\end{equation*}
%
Next we investigate uncertainty reduction criteria related to those indicators.
In order to be able to calculate them in semi-analytical forms relying on integrals of multivariate Gaussian cumulative 
distribution functions, we will appeal to some key intermediate results presented next.

>>>>

We now introduce quantities that allow to quantify the uncertainty on the volume of the excursion set $\es$. Let $\mes$ be a 
(locally finite, Borel) measure  on $\domain$. We want to investigate the probability distribution 
of $\mes(\es)$ through its moments.

@@

Centred moments may be computed using Proposition~\ref{propo1} developed in the appendix. 
In particular, it allows to write the excursion measure variance $\emv = \operatorname{Var}[\mes(\es)]$ as an integral of the excursion probability
\begin{equation*}
\begin{split}
\emv
&=\int_{\domain^2} \mathbb{P}\left(
\gp[u]\in T, \gp[v]\in T \right)
d\mes^{\otimes}(u, v)\\
&-\left( \int_{\domain} \mathbb{P}\left(\gp[u]\in T\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
which boils down in the excursion/sojourn case where $\T=(-\infty, t_1] \times
\dots \times (-\infty, t_{\no}]$ to
\begin{equation*}
\begin{split}
\emv
%\operatorname{Var}[\mes(\es)]
&=\int_{\domain^2}
\varPhi_{2\no}
\left(
(\bt, \bt); \mu((u,v)),
K((u,v),(u,v))
\right)
\
\mathrm{d}\mes^{\otimes} %\mes
%\productMeasure
(u,v)\\
&-\left( \int_{\domain} \varPhi_{\no}\left(\bt;\mu(u), K(u)\right) d\mes(u) \right)^2,
\end{split}
\end{equation*}
%
Note that like in the case of scalar-valued fields, this quantity requires to work out an integral over $\domain^2$. In 
contrast, and still like in the scalar-valued case, the Integrated Bernoulli Variance (IBV) of \cite{Bect.etal} involves 
solely an integral on $\domain$ and can be expanded in our present settings as follows
\begin{equation*}
\begin{split}
\operatorname{IBV} %(\es) %;\mes)
&=\int_{\domain}
\mathbb{P}\left(\gp[\uu]\in T\right)(1-\mathbb{P}\left(\gp[\uu]\in T\right))
d\mes(u) \\
&=\int_{\domain}
\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right)
-\left(\varPhi_{\no}\left(\bt;\mu(\uu), K(\uu)\right) \right)^2
\mathrm{d}\mes(u).
\end{split}
\end{equation*}
%
Next we investigate uncertainty reduction criteria related to those indicators.
